\input{../header.tex}

\section{Dirichlet Processes}

\subsection{Bayesian Historgram}
Consider the issue of estimating a density using a histogram. If one is \emph{given} a partitioning scheme $\xi_0 < \xi_1 < \ldots < \xi_K$ then the probability of a value is given by
\begin{align}
p(y) = \sum_k \one_{\xi_{k-1}<y<\xi_k} \frac{\theta_k}{\xi_k - \xi_{k-1}} \label{chap1:eqn:bayes-hist}
\end{align}
A natural choice for the prior distribution over $\vv{\theta}$ is the Dirichlet parameterised by a vector $\vv{\alpha}$. This parameterisation can be factorised into $\alpha \vv{u}$ where $\vv{u}$ is a vector on the K-1 simplex. Hence
\begin{align}
\vv{\theta} &\sim \dir{\alpha \vv{u}} &
\ex{\theta_k}{} & = u_k = \frac{\alpha u_k}{\sum_j \alpha u_j} &
\var{\theta_k}{} & = \oneover{\alpha + 1} u_k (1 - u_k) \label{chap1:eqn:dir-variance-etc}
\end{align}
From this it can be seen that $\alpha$ acts as a precision, or concentration, parameter. Higher values of $\alpha$ will prefer draws closer to $\vv{u}$.

There are two standard ways of taking a draw from a Dirichlet. One takes $K$ draws $\hat{\theta}_k \sim \gam{\alpha u_k}{\lambda}$ -- where $\lambda$ can be any value -- and then normalises them so that $\theta_k = \frac{\hat{\theta}_k}{\sum_j \hat{\theta}_j}$. 

A different approach is obtained by observing that for $\vv{\theta} \sim \dir{\alpha \vv{u}}$ the marginal distribution of a single component is $\theta_k \sim \betadist{\alpha u_k}{1- \alpha u_k}$. Consequently one can sample $\vv{\theta}$ by repeatedly sampling from Beta distributions
\begin{align}
V_k & \sim \betadist{\alpha u_k}{\alpha \sum_{j>k} u_j} &
\theta_k & = V_k \prod_{j > k}(1 - V_j) &
\theta_K = 1 - \sum_{j=1}^{K} \theta_j
\end{align}
This is known as a stick-breaking construction. We begin with a stick of unit length, then break off a portion given by a draw from a Dirichlet distribution. We then take the remainder and break a scaled portion from that. We continue till we've reached the K-th element, at which point the \emph{entirety} of the remaining stick length is assigned to $\theta_K$.

One might reasonably ask what happens in the limit as $K \rightarrow \infty$. If our stick-breaking construction were extended, we would end up with infinitismally small draws.

\bflong{Infinitely long Dirichlet Distribution}

\newcommand \sampspace { { \Omega } }
\newcommand \borel { { \mathcal{B} } }

\subsection{Dirichlet Process}
Consider the standard definition of a probability space $(\sampspace, \borel, P)$ where $\sampspace $ is the sample space, $\borel$ is a $\sigma$-algebra -- i.e. a collection of sets closed under complement, intersection and countably infinite union -- and $P$ is a probability measure $P : \Sigma \rightarrow [0,1]$. For a finite subset of sets $B_k$ drawn from $\borel$ we can assign probabilities to these sets, or ``bins" according to

\begin{align}
P(B_1), \ldots, P(B_K) & \sim \dir{\alpha P_0 (B_1), \ldots, P_0 (B_K)} \end{align}
This presentation differs to \eqref{chap1:eqn:bayes-hist} in that it does not specify how the probability mass $P(B_k)$ should be distributed across the values in $B_k$. The base measure $P_0$ is also defined in $(\sampspace, \borel)$ and can be any distribution, e.g. a Gaussian in a mixture model. Standard conjugacy results provide that

\begin{align}
\begin{split}
P(B_1), \ldots, &P(B_k) | x_1, \ldots x_n
= \\
& \dir{\alpha P_0(B_1) + \sum_i \one_{y_i \in B_1}P_0(B_1), \ldots, \sum_i \one_{y_i \in B_K} P_0(B_K)}
\end{split}
\end{align}

On its own, this does not define a complete prior on $P$ as it is conditioned on a specific partition $B_1, \ldots, B_K$. Generalising this model\bfshort{A reference would be really handy here!} to handles all possible partitions $B_1, \ldots, B_K$, for all possible values of $K$, gives rise to the Dirichlet process.

\begin{align}
P \sim \DP{\alpha P_0}
\ex{P(B_k)}{} &= P_0 (B_k) &
\var{P(B_k)}{} &= \oneover{1 + \alpha} P_0(B_k) (1 - P_0 (B_k) )
\end{align}
The Dirichlet Process can be expressed using a Beta distribution
\begin{align}
P(B_k) = \betadist{\alpha P_0(B_k)}{1 - \alpha P_0(B_k)}
\end{align}
and is conjugate to a multinomial likelihood:




\input{../footer.tex}