\input{../header.tex}

\section{Plan and Timetable for Remaining Work}

The plan for future work is:
\begin{description}
\item[July to September 2015] $\text{ } $\\
Work is currently underway on the academic paper recommendation algorithm mentioned in the future work section. A dataset, ACL, has already been processed and cleaned, and a basic implementation of the model, without features or low-rank covariances, has been created with some interim results. The results thus far are competitive with respect to RTM, but suggest that using a single distribution over $\Theta$ is not sufficient for good recommendations, and so needs to be permuted for the two use-cases. The remaining work to be done is
\begin{enumerate}
    \item Amend the algorithm in light of the results thus far, incorporating an extra parameter for paper recommendations to account for inherent paper popularity.
    \item Process a second database, such as CoraKB
    \item Write up a possble paper for for submission to AI Stats, if feasible, for the November 25, 2015 deadline
\end{enumerate}
\item[September 2015] $\text{ } $\\
\begin{enumerate}
    \item Enter `continuing research-student status.
\end{enumerate}
\item[October 2015 - December 2015] $\text{ } $\\
Return to the multi-task admixture model. 
\begin{enumerate}
    \item Investigate the use of sparsity, and see if it can give competitive results. 
    \item Implement a Dirichlet Multinomial Regression (DMR) baseline model. 
    \item Add an additional dataset, the current candidate is the IAPR TC-12 image labelling dataset for which the LEAR group at Inria, Grenoble, has extracted features and made them available at \url{http://lear.inrialpes.fr/people/guillaumin/data.php}.
    \item The AI-Stats deadline for submission of the recommendation project is November 25, 2015
\end{enumerate}
\item[January to March 2016] $\text{ }$\\
    \begin{enumerate}
        \item Write and submit a final thesis
        \item Submit the micro-text prediction work, if competitive, to SIGKDD. The deadline is March 6, 2016. 
    \end{enumerate}
\end{description}


\input{../footer.tex}