\input{../header.tex}

\subsubsection{Priors with Kronecker-Product Covariance Structure}
An alternative to employing a single multivariate Gaussian prior over many weight \emph{vectors} is to employ a single matrix-variate Gaussian\cite{Gupta1999} prior over a weight \emph{matrix} $W = \left\{\vv{w}_l\right\}_{l=1}^L$ where $W \in \MReal{L}{F}$. This is denoted as as $\mnor{M}{\Omega}{\Sigma}$ where $\Omega \in \MReal{F}{F}$ and $\Sigma \in \MReal{L}{L}$ are called the row and column covariances, $M \in \MReal{L}{F}$ is the mean matrix, and the density is defined as:
\begin{equation}
p(W) = (2\pi)^{KF}|\Sigma|^{-F}|\Omega|^{-K} \Etr{\half \inv{\Sigma}\left(W - M\right) \inv{\Omega} \left(W - M \right)\T}
\end{equation}

where $\Etr{X} = \exp (\Tr{X})$.

This is related to the multivariate normal distribution by the identity $\mnor{M}{\Omega}{\Sigma} = \nor{\vecf{M}}{\Sigma \otimes \Omega}$ where $\vecf{\cdot}$ is the function that converts a matrix to a vector by stacking its columns. 

The row and column covariances $\Omega$ and $\Sigma$ are unidentifiable as $A \otimes B = \lambda A \otimes \frac{1}{\lambda}B$ for any scalar $\lambda$. In practice this is resolved by forcing the one of the elements on the diagonals of the matrices, such as the the last, to 1 which gives rise to the iterative ``flip-flop" algorithm\cite{Srivastava2009} for maximum-likelihood estimation Given N observations of a matrix $X_n \in \MReal{L}{F}$, one can fit a matrix-variate Gaussian of the form $\mnor{M}{\alpha^2 \Omega}{\Sigma}$ using the following update

\begin{align}
\Omega & = \frac{1}{\hat{\Omega}}_{ll} \hat{\Omega}, \qquad\qquad\hat{\Omega} = \frac{1}{N F} \sum_n (\vv{x} - \vv{m}) \inv{\Sigma} (\vv{x} - \vv{m})\T \\
\Sigma & = \frac{1}{\hat{\Sigma}}_{ff} \hat{\Sigma}, \qquad\qquad\hat{\Sigma} = \frac{1}{N L} \sum_n (\vv{x} - \vv{m})\T \inv{\Omega} (\vv{x} - \vv{m}) \\
\alpha^2 & = \frac{1}{NFL} \sum_n (\vv{x} - \vv{m})\T\invb{\Sigma \otimes \Omega}(\vv{x} - \vv{m})
\end{align}
where $\vv{x} = \vecf{X}$, $\vv{m} = \vecf{M}$ and $M=\frac{1}{N} \sum_n X_n$. These updates are iteratively evaluated till convergence.

%Note that $\Omega \in \MReal{L}{L}$ but its estimator is scaled by $\frac{1}{N F}$ and similarly for $\Sigma$. 
It has been proven that if $N > \max(L, F)$ then this procedure should converge on unique estimates of $\Sigma$ and $\Omega$\cite{Srivastava2009}. Similarly, this approach can be extended\cite{Srivastava2009a} to the case of a matrix-variate distribution $\mnor{XZY}{\Omega}{\Sigma}$ where $X$ and $Y$ are known design matrices and $Z$ is a latent matrix, providing unique estimators of $Z$, $\Omega$ and $\Sigma$ under the identifiability condition that the last element of the diagonal in $\Omega$ be set to one.

%Further extensions have been derived separable covariances with Toeplitz\cite{Wirfalt2010} and per-symmetric\cite{Jansson} structures. 

The same algorithm can also be extended to more than two dimensions\cite{Ohlson2011}\cite{Ohlson2013}, creating a multilinear normal distribution over tensors by means of a multi-linear $\vecf{\cdot}$ function with particular emphasis on the ``doubly-separable" case where the covariance is factorized into three covariance matrices.

In the Bayesian case, the use of shrinkage priors, such as the Inverse-Wishart distribution, on the covariances penalizes explosions and implosions\cite{Archambeau2011} and so makes the covariances weakly identifiable.

Matrix-variate priors can be employed to derive efficient algorithms for multi-task regression. Consider a Gaussian process, typically\cite{Bishop2006} presented as:
\begin{align}
\vv{w} \sim & \nor{0}{\alpha^2 I_F} & \vv{t}|\vv{w},X & \sim \nor{X\vv{w}}{\sigma^2 I_N} \\
& & \vv{t} & \sim \nor{\vv{0}}{\sigma^2I_N + K}
\end{align}
where $K = \frac{1}{\alpha^2} X X\T$ and via the kernel trick\cite{Jst2004} can be replaced with any other valid kernel. To extend this to the multi-task case, the single weight vector can be replaced with a weight matrix $W$ and similarly the target vector $t$ is with a matrix $T \in \MReal{L}{F}$. Using the $\vecf{\cdot}$ function to represent these with multivariate distributions and employing standard Gaussian\cite{Bishop2006} and Kronecker\cite{Minka2000a} identities once can derive the following multi-task GP regression algorithm\cite{Bonilla2008}:

\begin{align}
\vecf{W} \sim & \nor{0}{\alpha^2 I_F \otimes I_L} & 
\vecf{T}|W,X & \sim \nor{\vecf{XW}}{\sigma^2 I_N \otimes I_L} \\
& & \vecf{T} & \sim \nor{\vv{0}}{\Sigma \otimes K + I_L \otimes \sigma^2I_N}
\end{align}
Despite employing a matrix-variate prior over functions, the marginal distribution over T is not equivalent to any matrix-variate distribution as the sum in the covariance precludes any factorisation into a Krnoecker product of two matrices. One could recover a matrix-variate distribution in the no-noise case (i.e. $\sigma^2 = 0$) but this leads to additional complications, discussed later.

The mean prediction $\vv{t}_l^*$ for the l-th task given features $\vv{x}^*$ follows the standard form for a GP.

\begin{align}
m(\vv{x}^*)_l & = \left(\vv{k}_l \otimes \vv{k}_x^* \right)\T\inv{C} \vv{t} & C & = \Sigma \otimes K + \sigma^2 I_N \otimes I_L
\end{align}

where $\vv{k}_l$ is the l-th column of the task covariance matrix $\Sigma$ and $\vv{k}_*$ is the usual vector of covariances between the input $\vv{x}_*$ and all other points.

This model has been independently derived in the geostatistics literature, where it is is considered a co-kriging method known as the intrinsic correlation model\cite{Wackernagel1998}. A counter-intuitive property of this model is that in the no-noise case -- where $\sigma^2 = 0$ and so the covariance of $\vecf{T}$ simplifies to $\Sigma \otimes K$) -- no transfer of learning occurs. 

The reason is that in the case maximum of likelihood estimation this leads to an update of $K$ proportional to $-L \ln |K| - N \ln |T\T \inv{K} T|$, an expression that does not involve $\Sigma$, whereas the update for $\Sigma$ is $\frac{1}{N}T\T \inv{K} T$. In short, the targets are correlated using $\Sigma$ and $K$, but we can decorrelate those targets so that only $\Sigma$ remains. This is clearer if we work through the equation for the mean of the predicted target $\vv{t}_* \in \VReal{L}$ for a new input $\vv{x}^*$ in the no noise case:

\begin{align}
m(\vv{x}_*) & = (\Sigma \otimes \vv{k}_*)\T\invb{\Sigma \otimes K}\vecf{T} \\
& = (\Sigma\T \otimes \vv{k}_*\T)\left(\inv{\Sigma} \otimes \inv{K}\right)\vecf{T}\\
& = \left(\Sigma \inv{\Sigma} \right) \otimes \left(\vv{k}_*\T \inv{K}\right)\vecf{T} \\
& = \left( \begin{array}{c}
     \vv{k}_*\T \inv{K} \vv{t}_{\cdot 1} \\
     \vdots \\
     \vv{k}_*\T \inv{K} \vv{t}_{\cdot L} \\
 \end{array}\right) \\
\end{align}

With regard to inference, in published work\cite{Bonilla2008} no flip-flop algorithm has been employed: presumably the use of a parameterised covariance function instead of a covariance matrix ensured that no identifiability issues arose in practice.

%With regard to inference, the authors did not use any of the tricks outlined earlier to handle identifiability: presumably the use of a parameterised covariance function instead of a covariance matrix ensured that no identifiability issues arose in practice, however they did factor the task covariance $\Sigma = L L\T$ according to the Cholesky decomposition to ensure that the resulting estimate of the covariance was always positive semi-definite. 

%Additionally, a number of approximations were used to make calculation of the kernel feasible, such as subselecting a sparse subset of inputs to calculate the kernel matrix\fixme{Nystrom}.

An alternative approach the method outlined above is to introduce a noise-free auxiliary variable, Y, such that $T|Y\sim\nor{\vecf{Y}}{\sigma^2 I_{LF}}$, $Y\sim\nor{\vv{0}}{\Sigma \otimes \Omega}$\cite{Stegle2011}. The advantage of this model is that $Y$ and $T$ therefore have a matrix-variate distributions, and so there's no need to resort to multi-variate forms as outlined above. 

Such a model could further be extended (\cite{Stegle2011} again) to encode a preference for sparse feature by following the Glasso approach\cite{Friedman2008} and putting a Laplace prior on $\vecf{\inv{\Sigma}}$, while representing $\Omega$ as a kernel which captures task correlations. The case of applying the Glasso approach to both covariances, in a parametric, maximum likelihood setting, has additionally been investigated by \cite{Tsiligkaridis2012b} and \cite{Tsiligkaridis2012}, while a much simpler approach of adding an L1 penalty term, for both matrices, to the log-likelihood has been presented in\cite{Zhang2010a}.

%\fixme{cite Glasso}

While the use of matrix-variate priors has been motivated here as a means of facilitating multi-task learning, such methods can also be used as a way of factorising a large matrix into two considerably smaller ones. (from $(LF)^2$ to $L^2 + F^2$) enabled by the Kronecker product structure.

For example, in the case of the matrix completion problem where rows or columns can be treated as independent features, a natural distribution over the matrix is as matrix-normal with separate covariances. As there is only one matrix, $X \in \MReal{D}{P}$ one cannot reliably estimate the mean matrix, but if the mean matrix where itself factored into row and column parts, one would arrive at the model\cite{Allen2010}  $X \sim \mnor{\vv{m}_c \one_P\T + \vv{m}_r \one_N\T}{\Omega}{\Sigma}$ where $\vv{m}_c \in \VReal{D}$ is the mean of the columns, and $\vv{m}_r \in \VReal{P}$ is the mean of the rows. This can equally be explained as a random-effects model with $X_{dp} = m_{c,d} + m_{r,p} + \epsilon_{dp}$ where $\epsilon_{dp} \sim \nor{0}{\Sigma_{dd}\Omega_{pp}}$. Unfortunately, the fact that typical recommender datasets features large number of both users and products limits the applicability of this approach unless one simultaneously reduces the rank of \emph{both} covariances.


%A non-parametric extension of this model was presented in\cite{Yu2009} where $y_{nl} = m_{nl} + f_{nl} + \epsilon_{nl}$, $m_{nl}$ is a function of the task and input features, $f_{nl}$ are the random effects and $\epsilon_{nl}$ is IID zero-mean Gaussian noise. A matrix-variate GP was employed for $m_{nl} \sim \gp{0}{\Omega_0 \otimes \Sigma}$ while $L$ multi-variate GPs were employed for the random effects $f_{l} \sim \gp{0}{\tau \Sigma}$. An inverse Wishart process\cite{Dawid1981} was used as a prior for $\Sigma \sim \mathcal{IWP} \left(\kappa, \Sigma_0 + \lambda\delta\right)$. $\Omega_0$ and $\Sigma_0$ were the covariance functions over task and input features respectively. The marginal distribution over Y is shown to be a matrix-variate Student-T process


\input{../footer.tex}
