\input{../header.tex}

\subsection{Multi-Task Learning}

There are many cases where one needs to make many predictions from the same data. Examples include predicting scores in many different exams for a school-child\cite{Bonilla2008}\cite{Evgeniou2005} or predicting affinity to different aspects of a product for a single customer\cite{Allenby1999}. As shown in \cite{Caruana1997} in the case where these prediction \emph{tasks} are related, performance can be improved, particularly when data is limited, by allowing a transfer of knowledge between tasks; however \cite{Caruana1997} also showed that simply assuming inter-task relatedness where no relationship exists may be detrimental.

The multitask setting usually consists of a dataset $\mathcal{S} = \left\{\left(\vv{x}_d, \vv{t}_d\right)\right\}_{d=1}^D$ where $\vv{x}_n \in \VReal{F}$ are the features and $\vv{t} \in \VReal{L}$ are the outputs for each task. In some cases, where not all task outputs are known for each datapoint, or where task-specific features exist, this presentation is amended to $\mathcal{S} = \left\{\left\{\left(\vv{x}_{dl}, t_l\right)\right\}_{l=1}^L\right\}_{d=1}^D$. 

As described in\cite{Argyriou2005} the method of information transfer broadly falls into two categories. In the first, one assumes that tasks have some common underlying representation, which can be shared\cite{Caruana1997}. In fact, Dirichlet Multinomial Regression\cite{Mimno2008} is an example of this: while presented as a method of learning mixture probabilities in an admixture of multinomials for text, it can equally be thought of as a multi-task learning problem, issuing predictions for individual words by employing a topic-model to learn a low-dimensional topic-space from which word-counts are generated. 

The other approach, and the one which will be the focus of this section, is that of assuming that task-specific functions are ``close" together according to some metric. In the case of optimisation-based approached to machine learning, a number of approaches have been investigated to see if appropriate norms could be constructed to transfer knowledge while either learning a low-dimensional feature-space\cite{argyriou2007spectral}; learning sparse features\cite{Argyriou2005}; or using kernels to capture correlations\cite{Evgeniou2005}

In the Bayesian case, this naturally corresponds to learning the prior over functions. A straightforward approach\cite{Allenby1999} is simply to place the same Gaussian prior over each task-specific weight vector $\vv{w}_l$. By learning the parameters of this prior, information can be transferred between weight vectors. This basic framework can then be extended by employing a \emph{mixture} of Gaussians as the prior over weights\cite{BakkerHeskes2003}, enabling tasks to be clustered, with information shared between clusters only. Such an approach handles the issue of task heterogeneity, where different subsets of tasks are related to each other in different ways.


%Learning Multiple Tasks with a Sparse Matrix-Normal Penalty\cite{Zhang2010a}
%
%Multi-task relationship learning\cite{Zhang2012mtrl}

\input{../footer.tex}
